---
title: "Analyzing Friends Dialogue"
author: "Haddi Sise"
execute: 
  warning: false
  message: false
---

<h1>Analyzing Friends Dialogue</h1>

This data was obtained from user EmilHvitfeldt on Github. You can find the link <a href=https://github.com/EmilHvitfeldt/friends/tree/master/data>here</a>. I will use the data to analyze how much they mention dating over the years of the show being aired and how viewership varies based on the name of the main character in the title of the episode.

```{r}
# load packages
library(tidyverse)
```

```{r}
# load data sets 
load(file="friendsdata/friends.rda")
load(file="friendsdata/friends_info.rda")
```

```{r}
# how many times is the word date, dating, or dated is said per episode depending on the year

basic_info <- friends_info |>
  mutate(year_aired = str_sub(air_date, 1,4))|>
  group_by(year_aired)|>
  mutate(numepyear = n()) |>
  select(season, episode, year_aired, numepyear)

friend_date <- friends |>
  left_join(basic_info, by = c("season", "episode")) |>
  group_by(year_aired) |>
  filter(str_detect(text, "(?i)\\bdate|dating\\b") == TRUE) |>
  summarise(date_ment = n(), numepyear)  |>
  distinct() |>
  mutate(datepeppyear = date_ment/numepyear)

ggplot(friend_date, aes(x = year_aired, y = datepeppyear)) +
  geom_col() +
  theme_bw() +
  labs(
    x = "Year Aired",
    y = "Number of times date said per episode",
    title = "Concept of Dating Mentioned Over the Years in Friends Dialogue"
  )
```

In 1995,1996, 1998, and 1999 the word date was said a lot less often in friends. I have never watched Friends but I assume that during these years, the main characters were in more long term relationships. This is probably why we see a dip in 2004 as well because they are wrapping up the show and giving the characters happy endings with a partner. Again this paragraph is written by someone who has never watched Friends.

```{r, fig.height=10}
# in the titles how common is it for a main character to get mentioned

main <- friends_info |>
  filter(str_detect(title,"The One") == TRUE) |>
  mutate(title = str_replace(title, "The One ", "The One:")) |>
  mutate(spec_title = str_extract(title, "(?<=:).+"))|>
  mutate(main_char = if_else(str_detect(spec_title, 
                                "(?i)\\bchandler|monica|joey|phoebe|rachel|ross\\b"), str_extract(spec_title, 
                                "(?i)\\bchandler|monica|joey|phoebe|rachel|ross\\b"), "None"))|>
  inner_join(friends, by = c("season", "episode")) |>
  select(season, episode, main_char, speaker) |>
  filter(str_detect(speaker, ("(Chandler Bing|Monica Geller|Joey Tribbiani|Phoebe Buffay|Rachel Green|Ross Geller)"))) |>
  filter(speaker != "Joey Tribbiani Sr.") |>
  group_by(main_char, speaker) |>
  summarize(numlines = n()) |>
  ungroup() |>
  group_by(main_char) |>
  mutate(percmainlines = (numlines/sum(numlines))*100)
  

ggplot(main, aes(fill = speaker, x = speaker, y = percmainlines)) +
  geom_col() +
  theme_bw() +
  labs(x = "Speaker",
       y = "Percentage of Main Character Lines",
       title = "Friends Episode Viewership Based on Main Character in Title") +
  facet_wrap(~main_char) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

Rachel consistently has a lot of lines compared to others in the cast even when she is not the main character based on the title of the episode. When a character has their name in the title, they  have the highest percentage of main character lines, except in the case of Monica.


