---
title: "PHD's Awarded in the US"
author: "Haddi Sise"
execute: 
  warning: false
  message: false
---

<h1>PhD's Awarded in the US</h1>

This data set was acquired from the National Center of Science and Engineering Statistics. Kelly Kang, a project officer compiled demographic information, educational history, sources of financial support, and post graduation plans of doctoral recipients. The link to this original data can be found <a href="https://ncses.nsf.gov/pubs/nsf19301/data">here</a>. Then I used a tidy data set from <a href="https://github.com/rfordatascience/tidytuesday/tree/main/data/2019/2019-02-19">github</a>. Using this data set, I chose to analyze which broad fields were awarding the most PhDs over the years.

```{r}

# loading necessary packages
library(tidyverse)
```

```{r}
# beginning data wrangling

#load csv files into data frame that I can work with
phds <- read.csv("phd_by_field.csv")

phds |>
  group_by(broad_field, year) |>
  filter(!is.na(n_phds)) |>
  summarize(tot_phds = sum(n_phds), year) |>
  distinct() |>
  ggplot(aes(x = year, y = tot_phds, color = broad_field)) +
    geom_smooth(se = FALSE) +
    theme_classic() +
    labs(
      x = "Year",
      y = "PHD's Awarded",
      title = "PHD's Awarded in Various Fields",
      color = "Broad Field of PHD"
    )
```

Life sciences awards the most PhDs by a healthy margin and it looks like the amount being awarded is still on an increasing trend. Engineering has the least PhDs awarded per year by broad field but it looks like the amount of engineering PhDs being awarded is increasing. Education PhD's were on a decline but it looks like they are beginning to level off around 5000 PhDs awarded every year. More psychology and social sciences PhDs are being awarded each year.
